### 2022-03-07 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Research Compendium Template

This is a research compendium template geared towards producing a manuscript output as a Word document reproducibly using the `targets` package for workflow management.
See the [`targets` documentation](https://docs.ropensci.org/targets/) for more information on how to program analyses in this style.
I've found that writing `paper.Rmd` using RStudio's [visual markdown editor](https://rstudio.github.io/visual-markdown-editing/) and automatic text wrapping (with each sentence as a new line) works best for version control and collaboration with git and GitHub.
The visual editor also integrates seamlessly with [Zotero](https://www.zotero.org/) for adding in-text citations and bibliographies.
The [`renv` package](https://rstudio.github.io/renv/articles/renv.html) my also be useful for collaboration to keep package version synced across collaborators.
Before submitting the manuscript, I like to [archive my GitHub repository with Zenodo](https://guides.github.com/activities/citable-code/) and add a DOI badge to the README, and a link to the Zenodo archive to the Data Availability Statement.
In the `review` directory, you'll find my preferred way of structuring response to reviewers complete with the ability to cross-reference responses to avoid repeating yourself.

**GithHub Actions** This template also uses [GitHub actions](https://github.com/features/actions) to automatically render a .html version of the manuscript every time an edit to docs/paper.Rmd is made.
If [GitHub pages](https://pages.github.com/) are active, then the rendered draft is viewable at \<username\>.github.io/\<repository name\>/paper.html.
The workflow for this lives in .github/workflows/build-manuscript.yaml and can be disabled on GitHub in the "Actions" tab.

Readme text for *your* repository below:

------------------------------------------------------------------------

<!-- badges: start -->

<!-- DOI badges go here -->

<!-- badges: end -->

This repository contains data and code for a manuscript by \_\_\_\_\_\_ written for submission to \_\_\_\_\_\_ tentatively titled: \_\_\_\_\_.

The most recent draft of the manuscript is available [here](https://aariq.github.io/paper-template/paper.html).

# Reproducibility

This research compendium has been developed using the statistical programming language R.
To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip file from from this URL: [master.zip](/archive/master.zip).

To run the compendium and reproduce all outputs:

-   Open the project in RStudio by double-clicking the `.Rproj` file.
-   Install packages listed in `packages.R` (if this compendium uses `renv`, this may happen automatically or with `renv::restore()`).
-   Run scripts in /R/ in numerical order.
-   Knit /doc/paper.Rmd to produce output

Additional instructions if project uses `targets`:

-   Install the [`targets` package](https://docs.ropensci.org/targets/).
-   Run `targets::tar_make()` or `targets::tar_make_clustermq()` from the R console to run all code and produce all outputs.

**NOTE**: the GitHub actions in .github/workflows/build-manuscript.yaml will not work with a `targets` project.
Consider `targets::tar_github_actions()` as an alternative