### 2022-03-07 README.Rmd
customize templates

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- badges: start -->

<!-- DOI badges go here -->

<!-- badges: end -->

This repository contains data and code for a manuscript by Eric Scott and Emilio Bruna written for submission to The American Naturalist tentatively titled: \_\_\_\_\_.

The most recent draft of the manuscript is available [here](https://brunalab.github.io/lagged-ipms/paper.html).

# Reproducibility

This research compendium has been developed using the statistical programming language R.
To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip file from from this URL: [master.zip](/archive/master.zip).

To run the compendium and reproduce all outputs:

-   Open the project in RStudio by double-clicking the `.Rproj` file.
-   Install packages listed in `packages.R` (if this compendium uses `renv`, this may happen automatically or with `renv::restore()`).
-   Run scripts in /R/ in numerical order.
-   Knit /doc/paper.Rmd to produce output

<!-- Additional instructions if project uses `targets`: -->

<!-- -   Install the [`targets` package](https://docs.ropensci.org/targets/). -->

<!-- -   Run `targets::tar_make()` or `targets::tar_make_clustermq()` from the R console to run all code and produce all outputs. -->

<!-- **NOTE**: the GitHub actions in .github/workflows/build-manuscript.yaml will not work with a `targets` project. -->

<!-- Consider `targets::tar_github_actions()` as an alternative -->