### 2022-04-21 notes/understanding stochastic lambda.Rmd
figuring out what's up with IPMs

---
title: "Why is Stochastic Lambda > Deterministic?"
author: "Eric R. Scott"
date: "2022-04-21"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    highlight: kate
    theme:
      version: 4
      bootswatch: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

Figure out what the heck is going on.

# Load Data

```{r include=FALSE}
## Load your packages, e.g. library(targets).
source(here("packages.R"))

## Load your R files
lapply(list.files(here("R"), full.names = TRUE), source)

withr::with_dir(here(),{
  tar_load(data_full)
  tar_load(vit_other_ff)
})
```

# Pick two years, worst and best

Fit determinsitic IPMs to all years separately to identify a bad year and a good year.
Then subset data to only contain those years

```{r warning=FALSE}
lambdas <- data_full %>%
  filter(year>1998) %>% 
  group_by(year_fct) %>% 
  group_split() %>% 
  map_dbl(~{
    vit_list_det_ff = c(
      list(
        vit_surv = surv_det(.x, habitat = "1-ha"),
        vit_size = size_det(.x, habitat = "1-ha"),
        vit_flwr = flwr_det(.x, habitat = "1-ha"),
        vit_size_sdlg = size_sdlg_det(.x, habitat = "1-ha"),
        vit_surv_sdlg = surv_sdlg_det(.x, habitat = "1-ha")
      ),
      vit_other_ff)

    make_proto_ipm_det(vit_list_det_ff) %>% 
      make_ipm(iterations = 1000,  #only needs 100 to converge
               normalize_pop_size = FALSE, # to run as PVA
               usr_funs = list(get_scat_params = get_scat_params)) %>% lambda()
  })
worst_year <- c(1999:max(data_full$year))[which.min(lambdas)]; min(lambdas)
best_year <- c(1999:max(data_full$year))[which.max(lambdas)]; max(lambdas)

two_yrs <- data_full %>% filter(year %in% c(best_year, worst_year))
```

# Fit IPMs

## Deterministic

Fit a deterministic IPM with just those two years of data

```{r}
vit_list_det_ff = c(
  list(
    vit_surv = surv_det(two_yrs, habitat = "1-ha"),
    vit_size = size_det(two_yrs, habitat = "1-ha"),
    vit_flwr = flwr_det(two_yrs, habitat = "1-ha"),
    vit_size_sdlg = size_sdlg_det(two_yrs, habitat = "1-ha"),
    vit_surv_sdlg = surv_sdlg_det(two_yrs, habitat = "1-ha")
  ),
  vit_other_ff)

ipm_det_ff = make_proto_ipm_det(vit_list_det_ff) %>% 
  make_ipm(iterations = 1000,  #only needs 100 to converge
           normalize_pop_size = FALSE, # to run as PVA
           usr_funs = list(get_scat_params = get_scat_params))
lambda(ipm_det_ff)
```

## Stochastic

The stochastic IPM requires re-writing the proto_ipm function because it's now using different parameter sets for year.

```{r}
## forets fragments
vit_list_stoch_ff = c(
  list(
    vit_surv = surv_raneff(two_yrs, habitat = "1-ha"),
    vit_size = size_raneff(two_yrs, habitat = "1-ha"),
    vit_flwr = flwr_raneff(two_yrs, habitat = "1-ha"),
    vit_size_sdlg = size_sdlg_raneff(two_yrs, habitat = "1-ha"),
    vit_surv_sdlg = surv_sdlg_raneff(two_yrs, habitat = "1-ha")
  ), 
  vit_other_ff)

proto_ipm_stoch_ff <- 
  init_ipm(
    sim_gen = "general",
    di_dd = "di",
    det_stoch = "stoch",
    kern_param = "kern"
  ) %>%

  # Define growth and survival kernel
  define_kernel(
    "P_yr", #_yr makes a different P kernel for each value of `yr` from par_set_indices
    formula = s_z * G_z1z * d_log_size,
    family  = "CC",
    # Probability of survival
    s_z = predict(vit_surv,
                  newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                  type = 'response'),
    # Growth
    G_z1z = dt.scaled(
      log_size_2,
      df   = size_nu,
      mean = size_mu,
      sd   = size_sd
    ),
    size_mu = predict(vit_size,
                      newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                      type = 'response'),
    size_nu = get_scat_params(vit_size)["nu"],
    size_sd = get_scat_params(vit_size)["sd"],

    data_list = vit_list_stoch_ff,
    states          = list(c('log_size')),
    uses_par_sets   = TRUE,
    par_set_indices = list(yr = c(2001, 2003)),
    evict_cor       = TRUE,
    evict_fun       = truncated_distributions("t.scaled", "G_z1z")
  ) %>%

  # Define fecundity and recruitment kernel
  define_kernel(
    "go_sdlg_yr",
    formula = p_f * fruits * seeds * g_e * d_log_size,
    family  = "CD",
    # probability of flowering
    p_f = predict(vit_flwr,
                  newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                  type = 'response'),
    fruits = predict(vit_fruits,
                     newdata = tibble(log_size_prev = log_size_1),
                     type = 'response'), 
    seeds = exp(coef(vit_seeds)[1]), #mean seeds per fruit
    g_e = vit_germ_est,

    data_list     = vit_list_stoch_ff,
    states        = list(c('log_size', 'sdlg')),
    uses_par_sets = TRUE,
    par_set_indices = list(yr = c(2001, 2003))
  ) %>%

  define_kernel(
    name = "stay_sdlg",
    formula = 0,
    family = "DD",
    states = list(c("sdlg")),
    evict_cor = FALSE
  ) %>%

  # Enter into post-seedling
  define_kernel(
    name = "leave_sdlg_yr",
    formula = s_sdlg * G_z2_sdlg,
    family = "DC",
    s_sdlg = predict(vit_surv_sdlg,
                     newdata = tibble(year_fct = yr),
                     type = 'response'), #intercept from intercept-only glm
    G_z2_sdlg = dt.scaled(
      x    = log_size_2,
      df   = size_sdlg_nu,
      mean = size_sdlg_mu,
      sd   = size_sdlg_sd
    ),
    size_sdlg_mu = coef(vit_size_sdlg)[1], #intercept from intercept-only glm
    size_sdlg_nu = get_scat_params(vit_size_sdlg)["nu"],
    size_sdlg_sd = get_scat_params(vit_size_sdlg)["sd"],

    data_list = vit_list_stoch_ff,
    states = list(c("sdlg", "log_size")),
    uses_par_sets = TRUE,
    par_set_indices = list(yr = c(2001, 2003)),
    evict_cor = TRUE,
    evict_fun = truncated_distributions("t.scaled", "G_z2_sdlg")
  ) %>%

  # define implementation with midpoint rule
  define_impl(make_impl_args_list(
    kernel_names = c("P_yr",        "go_sdlg_yr", "stay_sdlg", "leave_sdlg_yr"),
    int_rule = "midpoint",
    state_start  = c("log_size", "log_size", "sdlg",     "sdlg"),
    state_end    = c("log_size", "sdlg",     "sdlg",     "log_size")
  )) %>% 
  #define lower bound, upper bound, and number of meshpoints for log_size
  define_domains(log_size = c(0, 8.018296, 100)) %>%  
  #arbitrary starting population state
  define_pop_state(n_log_size = runif(100), n_sdlg = .1)
```

Iterate IPM

```{r}
ipm_stoch_ff <- 
  proto_ipm_stoch_ff %>%
  make_ipm(iterations = 1000,
           normalize_pop_size = FALSE, # to run as PVA
           usr_funs = list(get_scat_params = get_scat_params))
```

# Investigation

First, with just the worst and best years, is the stochastic lambda still higher than the deterministic?

```{r}
lambda(ipm_det_ff)
lambda(ipm_stoch_ff, log = FALSE)
```

Yes, stochastic lambda is still higher!

What about the averages for each year?
Are they consistent with a bad and good year?

```{r}
all_lams <- tibble(year = ipm_stoch_ff$env_seq,
       lambda = lambda(ipm_stoch_ff, type_lambda = "all")[,1])
all_lams %>% group_by(year) %>% summarize(mean(lambda)) %>% knitr::kable()
```

Yes, but both the good and bad years mean lambdas are *greater* than the deterministic lambda.

How do they compare at the level of vital rates?

```{r}
newdata <- expand.grid(
  log_size_prev = seq(min(two_yrs$log_size_prev, na.rm = TRUE), max(two_yrs$log_size_prev, na.rm = TRUE), length.out = 50),
  year_fct = as.character(c(2001, 2003))
)

safe_predict <- possibly(predict, NA)

det_fitted <- 
  vit_list_det_ff %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>%
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "deterministic")

stoch_fitted <- 
  vit_list_stoch_ff %>%
  map_df(~safe_predict(.x, newdata= newdata, type = "response")) %>%
  bind_cols(newdata) %>%
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_")

bind_rows(stoch_fitted, det_fitted) %>% 
  ggplot(aes(x = log_size_prev, y = .fitted, color = year_fct)) +
  geom_line() +
  facet_wrap(~vital_rate, scales = "free")

```

Hold on.
Those numbers for size_sdlg don't make sense!
The mean size a seedling establishes at should be more like 2

```{r}
two_yrs %>% filter(sdlg_prev == TRUE, !is.na(log_size), habitat== "1-ha") %>% select(-L, -spei_history) %>% pull(log_size) %>% hist()
data_full %>% filter(ha_id_number == "2") %>% View()
data_full %>% filter(sdlg_prev == TRUE, !is.na(log_size), habitat == "1-ha") %>% View()
```