### 2022-04-22 notes/index.Rmd
link to github *pages*

---
output: html_document
---

# Notes:

```{r eval=FALSE, include=FALSE, results='asis'}
base_url <- "https://brunalab.github.io/lagged-ipms/notes/"
# paste(base_url, list.files(here::here("notes"), pattern = "[^index].html"), sep = "")
```

[Exploring different ways of fitting random effects GAMs](https://brunalab.github.io/lagged-ipms/notes/environmental-stochasticity.html)

[Why is stochastic lambda \> deterministic?](https://brunalab.github.io/lagged-ipms/notes/understanding-stochastic-lambda.html)

[Optimizing the number of meshpoints](https://brunalab.github.io/lagged-ipms/notes/optimizing-meshpoints.html)