### 2022-04-26 notes/optimizing bootstrapping.Rmd
add notes on optimizing bootstrapping

---
title: "Optimizing Bootstrapping"
author: "Eric R. Scott"
date: "2022-04-26"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    highlight: kate
    theme:
      version: 4
      bootswatch: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(targets)

source(here("R", "bcpi.R"))

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

How many bootstraps is enough to estimate CIs around lambda?

# Load Data

```{r data, echo=TRUE}
withr::with_dir(here::here(), tar_load(c(
  ipm_det_cf,
  ipm_det_ff,
  ipm_stoch_cf,
  ipm_stoch_ff,
  lambda_bt_det_cf,
  lambda_bt_det_ff,
  lambda_bt_stoch_cf,
  lambda_bt_stoch_ff
)))

```

```{r}
est <- lambda(ipm_det_cf)
df <- seq_along(lambda_bt_det_cf) %>% 
  map_df(~bcpi(t0 = est, t = lambda_bt_det_cf[1:.x]), .id = "B") %>% 
  mutate(B = as.numeric(B))

ggplot(df,aes(x = B)) +
  geom_line(aes(y = lower), color = "red") +
  geom_line(aes(y = upper), color = "blue")
```

Looks like it does converge after like 200 bootstraps

```{r}
est <- lambda(ipm_det_ff)
df <- seq_along(lambda_bt_det_ff) %>% 
  map_df(~bcpi(t0 = est, t = lambda_bt_det_ff[1:.x]), .id = "B") %>% 
  mutate(B = as.numeric(B))

ggplot(df,aes(x = B)) +
  geom_line(aes(y = lower), color = "red") +
  geom_line(aes(y = upper), color = "blue")
```

```{r}
est <- lambda(ipm_stoch_cf, log = FALSE)
df <- seq_along(lambda_bt_stoch_cf) %>% 
  map_df(~bcpi(t0 = est, t = lambda_bt_stoch_cf[1:.x]), .id = "B") %>% 
  mutate(B = as.numeric(B))

ggplot(df,aes(x = B)) +
  geom_line(aes(y = lower), color = "red") +
  geom_line(aes(y = upper), color = "blue")
```

For stochastic it does take longer.
Why the sudden leaps for the upper limit?

```{r}
est <- lambda(ipm_stoch_ff, log = FALSE)
df <- seq_along(lambda_bt_stoch_ff) %>% 
  map_df(~bcpi(t0 = est, t = lambda_bt_stoch_ff[1:.x]), .id = "B") %>% 
  mutate(B = as.numeric(B))

ggplot(df,aes(x = B)) +
  geom_line(aes(y = lower), color = "red") +
  geom_line(aes(y = upper), color = "blue")
```