### 2022-04-26 notes/understanding stochastic lambda.Rmd
update notes on stochastic lambda

---
title: "Why is Stochastic Lambda > Deterministic?"
author: "Eric R. Scott"
date: "2022-04-21"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    highlight: kate
    code_folding: show
    theme:
      version: 4
      bootswatch: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
set.seed(123)
```

# TODO

-   Get code to run

-   Instead of deterministic IPM using different vital rates models, use the random effects models, but plug in `newdata = (log_size_prev = log_size, year = "newlevel")` to get mean fitted values.

-   add analysis with CF

-   Don't normalize population sizes and plot both CF and FF as PVA for both models

-   Try year as fixed effect

-   Read Rees & Ellner 2009.
    Consider either fixed effect of year + kernel sampling OR random effect of year + parameter sampling.
    Maybe no added value for random effect of year + kernel sampling??

Rees, Mark, and Stephen P. Ellner.
"Integral Projection Models for Populations in Temporally Varying Environments." Ecological Monographs 79, no. 4 (2009): 575--94.
<https://doi.org/10.1890/08-1474.1>.

*Last compiled: `r Sys.Date()`*

# Purpose

Figure out what the heck is going on.

# Load Data

```{r include=FALSE}
## Load your packages, e.g. library(targets).
source(here("packages.R"))

## Load your R files
lapply(list.files(here("R"), full.names = TRUE), source)

withr::with_dir(here(),{
  tar_load(c(
    data_full,
    pop_vec_ff,
    pop_vec_cf,
    vit_other_ff,
    vit_other_cf
  ))
})

#don't need matrix columns for this
data_full <- data_full %>% select(-L, -spei_history)
```

# Pick two years, worst and best

Fit determinsitic IPMs to all years separately to identify a bad year and a good year.
Then subset data to only contain those years

```{r warning=FALSE, class.source = 'fold-hide'}
single_years <-   
  data_full %>%
  filter(year>1999) %>% 
  group_by(year_fct) %>% 
  group_split()

#fit vital rates for FF  
vit_list_yr_ff <- 
  single_years %>% 
  map(~dplyr::filter(.x, habitat == "1-ha")) %>% 
  map(~{
    vit_list_det = c(
      list(
        vit_surv = surv_det(.x),
        vit_size = size_det(.x),
        vit_flwr = flwr_det(.x),
        vit_size_sdlg = size_sdlg_det(.x),
        vit_surv_sdlg = surv_sdlg_det(.x)
      ),
      vit_other_ff)
    })
# make IPMs for FF
ipms_yr_ff <-
  map(vit_list_yr_ff, ~{
    make_proto_ipm_det(.x, pop_vec = pop_vec_ff) %>% 
      make_ipm(iterations = 1000,  #only needs 100 to converge
               normalize_pop_size = FALSE, # to run as PVA
               usr_funs = list(get_scat_params = get_scat_params))
  })

# fit vital rates for CF
vit_list_yr_cf <- 
  single_years %>% 
  map(~dplyr::filter(.x, habitat == "CF")) %>% 
  map(~{
    vit_list_det = c(
      list(
        vit_surv = surv_det(.x),
        vit_size = size_det(.x),
        vit_flwr = flwr_det(.x),
        vit_size_sdlg = size_sdlg_det(.x),
        vit_surv_sdlg = surv_sdlg_det(.x)
      ),
      vit_other_cf)
    })

#make IPMs for CF
ipms_yr_cf <-
  map(vit_list_yr_cf, ~{
    make_proto_ipm_det(.x, pop_vec = pop_vec_cf) %>% 
      make_ipm(iterations = 1000,  #only needs 100 to converge
               normalize_pop_size = FALSE, # to run as PVA
               usr_funs = list(get_scat_params = get_scat_params))
  })

#lambdas
lambdas_ff <- ipms_yr_ff %>% map_dbl(lambda)
lambdas_cf <- ipms_yr_cf %>% map_dbl(lambda)
```

```{r warning=FALSE}
#worst year
(worst_year_ff <- c(2000:max(data_full$year))[which.min(lambdas_ff)]); min(lambdas_ff)
(worst_year_cf <- c(2000:max(data_full$year))[which.min(lambdas_cf)]); min(lambdas_cf)

#best year
(best_year_ff <- c(2000:max(data_full$year))[which.max(lambdas_ff)]); max(lambdas_ff)
(best_year_cf <- c(2000:max(data_full$year))[which.max(lambdas_cf)]); max(lambdas_cf)

plot(2000:max(data_full$year), lambdas_cf, type = "l", col = "red"); lines(x =2000:max(data_full$year), y = lambdas_ff, col = "blue", type = "l")
```

2003 was the worst year in both habitats.
2001 was a good year in both habitats (but not the best in CF)

```{r warning=FALSE}
#combine good and bad year
two_yrs <- data_full %>% filter(year %in% c(2003, 2001))
two_yrs_ff <- two_yrs %>% filter(habitat == "1-ha")
two_yrs_cf <- two_yrs %>% filter(habitat == "CF")
```

# Fit IPMs

## Deterministic

Fit a deterministic IPM with just those two years of data

Fragments:

```{r warning=FALSE, class.source = 'fold-hide'}
vit_list_det_ff = c(
  list(
    vit_surv = surv_det(two_yrs_ff),
    vit_size = size_det(two_yrs_ff),
    vit_flwr = flwr_det(two_yrs_ff),
    vit_size_sdlg = size_sdlg_det(two_yrs_ff),
    vit_surv_sdlg = surv_sdlg_det(two_yrs_ff)
  ),
  vit_other_ff)

ipm_det_ff = make_proto_ipm_det(vit_list_det_ff, pop_vec = pop_vec_ff) %>% 
  make_ipm(iterations = 1000,  #only needs 100 to converge
           normalize_pop_size = FALSE, # to run as PVA
           usr_funs = list(get_scat_params = get_scat_params))
# FF
lambda(ipm_det_ff)
```

CF:

```{r warning=FALSE, class.source = 'fold-hide'}
vit_list_det_cf = c(
  list(
    vit_surv = surv_det(two_yrs_cf),
    vit_size = size_det(two_yrs_cf),
    vit_flwr = flwr_det(two_yrs_cf),
    vit_size_sdlg = size_sdlg_det(two_yrs_cf),
    vit_surv_sdlg = surv_sdlg_det(two_yrs_cf)
  ),
  vit_other_cf)

ipm_det_cf = make_proto_ipm_det(vit_list_det_cf, pop_vec = pop_vec_cf) %>% 
  make_ipm(iterations = 1000,  #only needs 100 to converge
           normalize_pop_size = FALSE, # to run as PVA
           usr_funs = list(get_scat_params = get_scat_params))
lambda(ipm_det_cf)
```

## Stochastic

The stochastic IPM requires re-writing the `proto_ipm` because it's now using different parameter sets for year.
This now uses vital rates models where there is a random effect of year fit such that each year can have a different shape smoother (but not different wiggliness).

```{r warning=FALSE, class.source = 'fold-hide'}
## forests fragments
vit_list_stoch_ff = c(
  list(
    vit_surv = surv_raneff(two_yrs_ff),
    vit_size = size_raneff(two_yrs_ff),
    vit_flwr = flwr_raneff(two_yrs_ff),
    vit_size_sdlg = size_sdlg_raneff(two_yrs_ff),
    vit_surv_sdlg = surv_sdlg_raneff(two_yrs_ff)
  ), 
  vit_other_ff)

proto_ipm_stoch_ff <- 
  init_ipm(
    sim_gen = "general",
    di_dd = "di",
    det_stoch = "stoch",
    kern_param = "kern"
  ) %>%

  # Define growth and survival kernel
  define_kernel(
    "P_yr", #_yr makes a different P kernel for each value of `yr` from par_set_indices
    formula = s_z * G_z1z * d_log_size,
    family  = "CC",
    # Probability of survival
    s_z = predict(vit_surv,
                  newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                  type = 'response'),
    # Growth
    G_z1z = dt.scaled(
      log_size_2,
      df   = size_nu,
      mean = size_mu,
      sd   = size_sd
    ),
    size_mu = predict(vit_size,
                      newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                      type = 'response'),
    size_nu = get_scat_params(vit_size)["nu"],
    size_sd = get_scat_params(vit_size)["sd"],

    data_list = vit_list_stoch_ff,
    states          = list(c('log_size')),
    uses_par_sets   = TRUE,
    par_set_indices = list(yr = c(2001, 2003)),
    evict_cor       = TRUE,
    evict_fun       = truncated_distributions("t.scaled", "G_z1z")
  ) %>%

  # Define fecundity and recruitment kernel
  define_kernel(
    "go_sdlg_yr",
    formula = p_f * fruits * seeds * g_e * d_log_size,
    family  = "CD",
    # probability of flowering
    p_f = predict(vit_flwr,
                  newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                  type = 'response'),
    fruits = predict(vit_fruits,
                     newdata = tibble(log_size_prev = log_size_1),
                     type = 'response'), 
    seeds = exp(coef(vit_seeds)[1]), #mean seeds per fruit
    g_e = vit_germ_est,

    data_list     = vit_list_stoch_ff,
    states        = list(c('log_size', 'sdlg')),
    uses_par_sets = TRUE,
    par_set_indices = list(yr = c(2001, 2003))
  ) %>%

  define_kernel(
    name = "stay_sdlg",
    formula = 0,
    family = "DD",
    states = list(c("sdlg")),
    evict_cor = FALSE
  ) %>%

  # Enter into post-seedling
  define_kernel(
    name = "leave_sdlg_yr",
    formula = s_sdlg * G_z2_sdlg,
    family = "DC",
    s_sdlg = predict(vit_surv_sdlg,
                     newdata = tibble(year_fct = yr),
                     type = 'response'), #intercept from intercept-only glm
    G_z2_sdlg = dt.scaled(
      x    = log_size_2,
      df   = size_sdlg_nu,
      mean = size_sdlg_mu,
      sd   = size_sdlg_sd
    ),
    size_sdlg_mu = coef(vit_size_sdlg)[1], #intercept from intercept-only glm
    size_sdlg_nu = get_scat_params(vit_size_sdlg)["nu"],
    size_sdlg_sd = get_scat_params(vit_size_sdlg)["sd"],

    data_list = vit_list_stoch_ff,
    states = list(c("sdlg", "log_size")),
    uses_par_sets = TRUE,
    par_set_indices = list(yr = c(2001, 2003)),
    evict_cor = TRUE,
    evict_fun = truncated_distributions("t.scaled", "G_z2_sdlg")
  ) %>%

  # define implementation with midpoint rule
  define_impl(make_impl_args_list(
    kernel_names = c("P_yr",        "go_sdlg_yr", "stay_sdlg", "leave_sdlg_yr"),
    int_rule = "midpoint",
    state_start  = c("log_size", "log_size", "sdlg",     "sdlg"),
    state_end    = c("log_size", "sdlg",     "sdlg",     "log_size")
  )) %>% 
  #define lower bound, upper bound, and number of meshpoints for log_size
  define_domains(log_size = c(0, 8.018296, 100)) %>%  
  #arbitrary starting population state
  define_pop_state(pop_vectors = pop_vec_ff)
```

```{r warning=FALSE, class.source = 'fold-hide'}
## forests fragments
vit_list_stoch_cf = c(
  list(
    vit_surv = surv_raneff(two_yrs_cf),
    vit_size = size_raneff(two_yrs_cf),
    vit_flwr = flwr_raneff(two_yrs_cf),
    vit_size_sdlg = size_sdlg_raneff(two_yrs_cf),
    vit_surv_sdlg = surv_sdlg_raneff(two_yrs_cf)
  ), 
  vit_other_cf)

proto_ipm_stoch_cf <- 
  init_ipm(
    sim_gen = "general",
    di_dd = "di",
    det_stoch = "stoch",
    kern_param = "kern"
  ) %>%

  # Define growth and survival kernel
  define_kernel(
    "P_yr", #_yr makes a different P kernel for each value of `yr` from par_set_indices
    formula = s_z * G_z1z * d_log_size,
    family  = "CC",
    # Probability of survival
    s_z = predict(vit_surv,
                  newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                  type = 'response'),
    # Growth
    G_z1z = dt.scaled(
      log_size_2,
      df   = size_nu,
      mean = size_mu,
      sd   = size_sd
    ),
    size_mu = predict(vit_size,
                      newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                      type = 'response'),
    size_nu = get_scat_params(vit_size)["nu"],
    size_sd = get_scat_params(vit_size)["sd"],

    data_list = vit_list_stoch_cf,
    states          = list(c('log_size')),
    uses_par_sets   = TRUE,
    par_set_indices = list(yr = c(2001, 2003)),
    evict_cor       = TRUE,
    evict_fun       = truncated_distributions("t.scaled", "G_z1z")
  ) %>%

  # Define fecundity and recruitment kernel
  define_kernel(
    "go_sdlg_yr",
    formula = p_f * fruits * seeds * g_e * d_log_size,
    family  = "CD",
    # probability of flowering
    p_f = predict(vit_flwr,
                  newdata = tibble(log_size_prev = log_size_1, year_fct = yr),
                  type = 'response'),
    fruits = predict(vit_fruits,
                     newdata = tibble(log_size_prev = log_size_1),
                     type = 'response'), 
    seeds = exp(coef(vit_seeds)[1]), #mean seeds per fruit
    g_e = vit_germ_est,

    data_list     = vit_list_stoch_cf,
    states        = list(c('log_size', 'sdlg')),
    uses_par_sets = TRUE,
    par_set_indices = list(yr = c(2001, 2003))
  ) %>%

  define_kernel(
    name = "stay_sdlg",
    formula = 0,
    family = "DD",
    states = list(c("sdlg")),
    evict_cor = FALSE
  ) %>%

  # Enter into post-seedling
  define_kernel(
    name = "leave_sdlg_yr",
    formula = s_sdlg * G_z2_sdlg,
    family = "DC",
    s_sdlg = predict(vit_surv_sdlg,
                     newdata = tibble(year_fct = yr),
                     type = 'response'), #intercept from intercept-only glm
    G_z2_sdlg = dt.scaled(
      x    = log_size_2,
      df   = size_sdlg_nu,
      mean = size_sdlg_mu,
      sd   = size_sdlg_sd
    ),
    size_sdlg_mu = coef(vit_size_sdlg)[1], #intercept from intercept-only glm
    size_sdlg_nu = get_scat_params(vit_size_sdlg)["nu"],
    size_sdlg_sd = get_scat_params(vit_size_sdlg)["sd"],

    data_list = vit_list_stoch_cf,
    states = list(c("sdlg", "log_size")),
    uses_par_sets = TRUE,
    par_set_indices = list(yr = c(2001, 2003)),
    evict_cor = TRUE,
    evict_fun = truncated_distributions("t.scaled", "G_z2_sdlg")
  ) %>%

  # define implementation with midpoint rule
  define_impl(make_impl_args_list(
    kernel_names = c("P_yr",        "go_sdlg_yr", "stay_sdlg", "leave_sdlg_yr"),
    int_rule = "midpoint",
    state_start  = c("log_size", "log_size", "sdlg",     "sdlg"),
    state_end    = c("log_size", "sdlg",     "sdlg",     "log_size")
  )) %>% 
  #define lower bound, upper bound, and number of meshpoints for log_size
  define_domains(log_size = c(0, 8.018296, 100)) %>%  
  #arbitrary starting population state
  define_pop_state(pop_vectors = pop_vec_cf)
```

Iterate IPM with 1000 iterations

```{r warning=FALSE}
ipm_stoch_ff <- 
  proto_ipm_stoch_ff %>%
  make_ipm(iterations = 1000,
           normalize_pop_size = FALSE, # to run as PVA
           usr_funs = list(get_scat_params = get_scat_params))

ipm_stoch_cf <- 
  proto_ipm_stoch_cf %>%
  make_ipm(iterations = 1000,
           normalize_pop_size = FALSE, # to run as PVA
           usr_funs = list(get_scat_params = get_scat_params))
```

# Investigation

First, with just the worst and best years, is the stochastic lambda still higher than the deterministic?

```{r}
all_lams <-
  list(ff = ipm_stoch_ff, cf = ipm_stoch_cf) %>% 
  map_df(
    ~tibble(year = .x$env_seq,
         lambda = lambda(.x, type_lambda = "all")[,1])%>% 
  #remove burnin
  slice(101:n()),
  .id = "habitat"
  )

```

```{r lambdas}
tribble(~model, ~habitat, ~year, ~lambda,
        "det", "ff", "both combined", lambda(ipm_det_ff),
        "stoch", "ff", "both combined", lambda(ipm_stoch_ff, log = FALSE),
        "det", "cf", "both combined", lambda(ipm_det_cf),
        "stoch", "cf", "both combined", lambda(ipm_stoch_ff, log = FALSE),
        "det", "ff", "2001", lambdas_ff[2],
        "det", "ff", "2003", lambdas_ff[4],
        "det", "cf", "2001", lambdas_cf[2],
        "det", "cf", "2003", lambdas_cf[4]
) %>% 
  bind_rows(
    all_lams %>%
      group_by(habitat, year) %>%
      summarize(lambda = mean(lambda)) %>% 
      mutate(year = paste(year, "mean"),
             model = "stoch")
  ) %>% arrange(model, habitat)
```

Stochastic lambda is still slightly higher than deterministic

Within stochastic IPMs, the mean lambdas for individual transition years are consistent with good and bad years, but both are *higher* than deterministic lambda with both years combined!

The deterministic lambda for 2001 is a lot higher than the mean of the 2001 lambdas from the stochastic simulation and the deterministic lambda for 2003 is a lot lower than the mean of the 2003 lambdas from the stochastic simulation.
This makes sense to me, as fitting year as a random effect rather than a fixed effect will shrink estimates toward a global mean.

How do they compare at the level of vital rates?

```{r class.source = 'fold-hide'}
newdata <- expand.grid(
  log_size_prev = seq(min(two_yrs$log_size_prev, na.rm = TRUE), max(two_yrs$log_size_prev, na.rm = TRUE), length.out = 50),
  year_fct = as.character(c(2001, 2003))
)

safe_predict <- possibly(predict, NA)

#deterministic, just 2001
det_fitted_2001_ff <- 
  vit_list_yr_ff[[2]] %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>% 
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "det. 2001")

#deterministic, just 2003
det_fitted_2003_ff <- 
  vit_list_yr_ff[[4]] %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>% 
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "det. 2003")

#deterministic with both years combined
det_fitted_ff <- 
  vit_list_det_ff %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>%
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "det. both years")

#stochastic
stoch_fitted_ff <- 
  vit_list_stoch_ff %>%
  map_df(~safe_predict(.x, newdata= newdata, type = "response")) %>%
  bind_cols(newdata) %>%
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = paste("stoch.", year_fct))

bind_rows(stoch_fitted_ff, det_fitted_ff, det_fitted_2001_ff, det_fitted_2003_ff) %>% 
  ggplot(aes(x = log_size_prev, y = .fitted, color = year_fct, linetype = year_fct)) +
  geom_line(alpha = 0.9) +
  facet_wrap(~vital_rate, scales = "free") +
  labs(title = "Forest Fragments")

```

```{r class.source = 'fold-hide'}
newdata <- expand.grid(
  log_size_prev = seq(min(two_yrs$log_size_prev, na.rm = TRUE), max(two_yrs$log_size_prev, na.rm = TRUE), length.out = 50),
  year_fct = as.character(c(2001, 2003))
)

safe_predict <- possibly(predict, NA)

#deterministic, just 2001
det_fitted_2001_cf <- 
  vit_list_yr_cf[[2]] %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>% 
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "det. 2001")

#deterministic, just 2003
det_fitted_2003_cf <- 
  vit_list_yr_cf[[4]] %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>% 
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "det. 2003")

#deterministic with both years combined
det_fitted_cf <- 
  vit_list_det_cf %>% 
  map_df(~safe_predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata) %>%
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = "det. both years")

#stochastic
stoch_fitted_cf <- 
  vit_list_stoch_cf %>%
  map_df(~safe_predict(.x, newdata= newdata, type = "response")) %>%
  bind_cols(newdata) %>%
  select(-vit_fruits, -vit_seeds, -vit_germ_est) %>% 
  pivot_longer(starts_with("vit_"), names_to = "vital_rate", values_to = ".fitted", names_prefix = "vit_") %>% 
  mutate(year_fct = paste("stoch.", year_fct))

bind_rows(stoch_fitted_cf, det_fitted_cf, det_fitted_2001_cf, det_fitted_2003_cf) %>% 
  ggplot(aes(x = log_size_prev, y = .fitted, color = year_fct, linetype = year_fct)) +
  geom_line(alpha = 0.9) +
  facet_wrap(~vital_rate, scales = "free") +
  labs(title = "Continuous Forest")

```

For the stochastic models, the two years are nearly indistinguishable for seedlings.
I believe this is because of shrinkage---there is little information for either year so means are shrunk toward global average.
You can see this more clearly for flowering probability where the order of lines at the larger end is 2001 deterministic (essentially a fixed effect), 2001 stochastic (random effect), both years combined, 2003 stochastic (random effect), and 2003 deterministic (fixed effect).