### 2022-05-10 README.Rmd
update README with roadmap (#18)

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- badges: start -->

<!-- DOI badges go here -->

<!-- badges: end -->

This repository contains data and code for a manuscript by Eric Scott and Emilio Bruna written for submission to The American Naturalist tentatively titled: \_\_\_\_\_.

The idea here is to investigate the consequences of including lagged effects (such as those modeled by distributed lag non-linear models) into population models (integral projection models, IPMs).
We'll compare population growth rates (lambda) with deterministic, stochastic (matrix shuffling), and lagged effects IPMs.

The most recent draft of the manuscript is available [here](https://brunalab.github.io/lagged-ipms/paper.html).

[Notes for collaborators](https://brunalab.github.io/lagged-ipms/notes/index.html)

# Roadmap

Before June 3:

-   [x] Finish building all IPMs
-   [x] Document functions and \_targets.R
-   [ ] Calculate bootstrapped 95% CIs around all lambdas
-   [ ] Finish writing methods section
-   [ ] Draft tables & figures
-   [ ] Archive draft of code in Zenodo

Future:

-   [ ] Finish writing and revising
-   [ ] Submit

# Reproducibility

This research compendium has been developed using the statistical programming language R.
To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip file from from this URL: [master.zip](/archive/master.zip).

To run the compendium and reproduce all outputs:

-   Open the project in RStudio by double-clicking the `.Rproj` file.
-   Install packages listed in `packages.R` (if this compendium uses `renv`, this may happen automatically or with `renv::restore()`).
-   Install the [`targets` package](https://docs.ropensci.org/targets/).
-   Run `targets::tar_make()` or `targets::tar_make_clustermq()` from the R console to run all code and produce all outputs.