### 2022-05-24 notes/low-density-continuous.Rmd
do some analysis to see if low and high density CF plots differ

---
title: "Is plot density important?"
author: "Eric R. Scott"
date: "2022-05-24"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    highlight: kate
    theme:
      version: 4
      bootswatch: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(targets)
library(mgcv)
library(gratia)
library(ipmr)
library(marginaleffects)

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

*Last compiled: `r Sys.Date()`*

# Purpose

The continuous forest plots are not sampled randomly.
The first three plots (CF-1--3) have a much higher density of plants than the 1-ha fragments.
Emilio then purposefully saught out low-density continuous forest plots with numbers of plants comparable to the plots in 1-ha fragments (CF-4--6).
Some preliminary investigation leads me to believe the high and low density plots may be different in terms of observed population growth, vital rates, and lambda.
I've fit the vital rates models and ran the IPMs using only low-density CF plots for comparison.

# Load Data

```{r data, echo=TRUE}
withr::with_dir(here::here(), {
  tar_load(c(
    lambda_table,
    ipm_det_cf_low,
    ipm_stoch_cf_low,
    ipm_dlnm_cf_low,
    data_full
  ))
})
```

# Vital rates

One simple way to look for differences is to fit vital rates models with different smooths for 1-ha, CF-low density, and cf-high density.
The `by = hab_dens` argument in `s()` generates a different smooth for each level of habitat/density

#### NOTE:

One important confounding factor here is that there is missing data for the year 2000 for all low-density CF plots and missing data for 2003 for one of the plots.
So differences between low and high density or between low density and 1-ha could be due to the year 2000.
To be "fair" I'm going to filter out 2000 (and 2001, since we need the size in the previous year) for all plots.

```{r}
# data_full$habitat %>% unique()
df <-
  data_full %>% 
  select(-L, -spei_history) %>% 
  filter(year != 2000) %>% 
  mutate(hab_dens = case_when(
    habitat == "1-ha" ~ "1-ha",
    plot %in% c("CF-1", "CF-2", "CF-3") ~ "CF-high",
    plot %in% c("CF-4", "CF-5", "CF-6") ~ "CF-low",
    TRUE ~ NA_character_
  ), .after = habitat) %>% 
  mutate(hab_dens = as.factor(hab_dens))

```

## Survival

Fit the model with different intercepts and smooths for the three habitat:density combos

```{r}
m_surv <- bam(
    surv ~ hab_dens + s(log_size_prev, bs = "cr", k = 20, by = hab_dens),
    family = binomial,
    data = df %>% filter(sdlg_prev == FALSE),
    method = "fREML"
)
draw(m_surv)
```

Look at the pairwise differences in these smooths:

```{r}
diff_surv <- difference_smooths(m_surv, "s(log_size_prev)")
ggplot(diff_surv, aes(x = log_size_prev, y = diff)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  geom_line() +
  facet_wrap(level_1 ~ level_2) +
  labs(title = "Difference smooths for survival")
```

Calculate contrast, the mean difference between adjusted predictions (See [vignette](https://vincentarelbundock.github.io/marginaleffects/articles/mfx02_contrasts.html)).
The important thing to get about how these work is they evaluate the model at values in the actual dataset, then average the differences.

```{r}
cmp_surv <- comparisons(m_surv, variables = list(hab_dens = "pairwise"))
summary(cmp_surv)
```

The mean difference in survival between low and high density CF is significant, and the mean difference between low density CF and 1-ha is not significant.

## Size

Fit the model with different intercepts and smooths for the three habitat:density combos

```{r}
m_size <- bam(
    log_size ~ hab_dens + s(log_size_prev, bs = "cr", k = 20, by = hab_dens),
    family = scat,
    data = df %>% filter(sdlg_prev == FALSE, surv == 1, !is.na(log_size)),
    method = "fREML"
)
draw(m_size)
```

Look at the pairwise differences in these smooths:

```{r}
diff_size <- difference_smooths(m_size, "s(log_size_prev)")
ggplot(diff_size, aes(x = log_size_prev, y = diff)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  geom_line() +
  facet_wrap(level_1 ~ level_2) +
  labs(title = "Difference smooths for size")
```

differences are very minor

Contrasts:

```{r}
cmp_size <- comparisons(m_size, variables = list(hab_dens = "pairwise"))
summary(cmp_size)
```

Here the CF high and 1-ha are not significantly different but other comparisons are.
The smooths for CF-high and 1-ha look really different for small plants, but because there are few small plants the comparison above shows a small effect.

## Flowering

```{r}
m_flwr <- bam(
    flwr ~ hab_dens + s(log_size_prev, bs = "cr", k = 20, by = hab_dens),
    family = binomial, 
    data = df %>% filter(sdlg == FALSE, surv == 1, !is.na(log_size)),
    method = "fREML"
  )
draw(m_flwr)
```

The smooth for high density CF looks differen than the other two.

Look at the pairwise differences in these smooths:

```{r}
diff_flwr <- difference_smooths(m_flwr, "s(log_size_prev)")
ggplot(diff_flwr, aes(x = log_size_prev, y = diff)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  geom_line() +
  facet_wrap(level_1 ~ level_2) +
  labs(title = "Difference smooths for size")
```

Contrasts:

```{r}
cmp_flwr <- comparisons(m_flwr, variables = list(hab_dens = "pairwise"))
summary(cmp_flwr)
```

all pairwise comparisons are significantly different and the biggest difference is actually between CF-low and 1-ha

#IPMs

We can also look at the level of the IPMs

```{r}
lambda_table %>% 
  add_row(ipm = "det", habitat = "cf_low", est = lambda(ipm_det_cf_low)) %>% 
  add_row(ipm = "stoch", habitat = "cf_low", est = lambda(ipm_stoch_cf_low, log = FALSE)) %>% 
  add_row(ipm = "dlnm", habitat = "cf_low", est = lambda(ipm_dlnm_cf_low, log = FALSE)) %>% 
  add_row(ipm = "dlnm", habitat = "ff", est = lambda(tar_read(ipm_dlnm_ff), log = FALSE)) %>% 
  add_row(ipm = "dlnm", habitat = "cf", est = lambda(tar_read(ipm_dlnm_cf), log = FALSE)) %>% 
  arrange(ipm, habitat)
```

For the deterministic and stochastic, kernel-resampled (random effect of year) models, CF low density plots have the lowest lambda.
For the DLNM IPM, CF and CF low density are comparable and FF has the lowest lambda.
The caveat is that the stochastic, kernel-resampled IPM for cf_low is iterated using a different sequence of years since it doesn't have data for 2000 or 2001.